---
title: "Project Proposal"
author: "Christophe"
date: "October 17, 2015"
output: 
  pdf_document:
        includes:
          in_header: header.tex
---

```{r include=FALSE, cache=FALSE}
# DO NOT REMOVE
# THIS IS FOR SETTING SOME PLOTTING PARAMETERS SO THAT YOUR PLOTS DON'T TAKE UP TOO MUCH SPACE
# IF YOU WOULD LIKE TO CHANGE THESE, SEE HELP FILES ON THE par() FUNCTION
# OR ASK FOR HELP
library(knitr)
## set global chunk options
opts_chunk$set(fig.path='figure/manual-', cache.path='cache/manual-', fig.align='center', fig.show='hold', par=TRUE)
## tune details of base graphics (http://yihui.name/knitr/hooks)
knit_hooks$set(par=function(before, options, envir){
if (before && options$fig.show!='none') par(mar=c(4,4,.2,.1),cex.lab=.95,cex.axis=.9,mgp=c(2,.7,0),tcl=-.3)
})
```

```{r pre-load, eval=TRUE, include=FALSE}
library(RCurl)
library(psych)
library(knitr)
url <- "https://raw.githubusercontent.com/ChristopheHunt/IS606_Fall_2015/master/Project/Data/"
```

```{r load_data, results = 'asis'}
# load data
library(RCurl)
url <- paste(url,"STD_Rate_Data_File.csv", sep = '')
std.data <- getURL(url)                
std.data <- read.csv(textConnection(std.data))
df <- data.frame(std.data)
df <- subset(df, df$Race.Ethnicity != 'Unknown' & df$Gender.Code != 'U')
df$STD.Cases <- as.numeric(df$STD.Cases)
df$Rate <- as.numeric(df$Rate)
display_variables <- c("Disease", "State", "Gender",
                       "Race.Ethnicity", "STD.Cases","Population","Rate")

display_data <- df[display_variables]
kable(head(display_data, n = 15))
```

### Research question 

Are there disparities in Chlamydia incidences when accounting for ethnicity and state/region in the US in 2013?

\pagebreak[4]

### Cases 

The number of cases and disease incidence rates are reported by year, gender of patient, race/ethnicity, type of STD, and state. 

To investigate this research question we are only taking cases reported for the year 2013. 

```{r cases, results = 'asis', eval=TRUE}
cases_table <- aggregate(list('Cases in 2013' = df$STD.Cases),
                by = list(Ethnicity = df$Race.Ethnicity, Gender = df$Gender), FUN=sum)
cases_table <- cases_table[order(cases_table$Ethnicity),]
kable(cases_table)

kable(aggregate(list('Cases in 2013' = df$STD.Cases), 
                by = list(State = df$State), FUN=sum))
```

### Data collection 

Beginning in 2003, all 50 states and the District of Columbia had converted from summary hardcopy reporting to electronic case-specific reporting via NETSS (with the exception of congenital syphilis, which is still reported on hardcopy forms by several areas). Also, before 1996, Chlamydia reporting was voluntary, and thus sporadic. From 1995 - 2000, upstate New York did not report Chlamydia, and thus the total US denominator population for Chlamydia excludes the New York state population for the years 1995 - 2000. However, New York City did report Chlamydia before the year 2000, and the New York City population and Chlamydia cases are included for 1984 - 2000. ([Sexually Transmitted Disease Morbidity by Race and Age.](http://wonder.cdc.gov/wonder/help/std-race-age-frame.html) (n.d.). Retrieved October 17, 2015.)

It is unclear how many layers of abstraction exists, it appears that the information is compiled from provider level reports of each patient, then aggregated by the various health agencies and reported to CDC. However, it could be the case that each patient or incidence is reported to the CDC and the CDC does the aggregation.  

### Type of study 

This study is an observational study on Sexually Transmitted Disease (STD) morbidity case records reported to the National Center for HIV/AIDS, Viral Hepatitis, STD, and TB Prevention (NCHHSTP), Centers for Disease Control and Prevention (CDC). 

\blandscape
\pagebreak[4]
  \global\pdfpageattr\expandafter{\the\pdfpageattr/Rotate 90} <!-- provided by http://tex.stackexchange.com/questions/40683/rotate-single-pdf-page-when-viewing --> 
  
### Data Source 

US Department of Health and Human Services, Centers for Disease Control and Prevention, National Center for HIV, STD and TB Prevention (NCHSTP), Division of STD/HIV Prevention, Sexually Transmitted Disease Morbidity 1996 - 2013, by gender, age group and race/ethnicity, CDC WONDER Online Database. 

CDC WONDER Online Database Query Date: Oct 17, 2015 8:30:27 AM


### Response 

The response variable or our variable of interest is the rate of STD (specifically Chlamydia). The rate of STD cases is measured as a numerical variable and at incidence per 1,000 people. 

### Explanatory 

There are three explanatory variables in the data set, they are the ethnicity, gender, and state. Ethnicity, gender, and state are considered to be categorical variables. 

### Relevant summary statistics 

Note that the sample size is 51 for all ethnicities because the observations are for all 50 states and the District of Columbia for 2013. 

```{r summary_stats, eval=TRUE, results='asis', tidy=TRUE}
sum_stat_m <- aggregate(list('Rate' = df$Rate),
                by = list(Ethnicity = df$Race.Ethnicity, State = df$State), FUN=sum)

sum_stat_m <- describeBy(sum_stat_m$Rate, sum_stat_m$Ethnicity, mat = TRUE)
rownames(sum_stat_m) <- sum_stat_m$group1
sum_stat_m <- sum_stat_m[-(5),-(1:3)]
sum_stat_m <- round(sum_stat_m,2)
kable(sum_stat_m, align = 'c')
```

\elandscape



