---
title: "Homework 9"
author: "Christophe Hunt"
date: "April 1, 2017"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
always_allow_html: yes  
---

\newpage

# Page 385: problem 1 a

Using the definition provided for the movement diagram, determine whether the following zero-sum games have a pure strategy Nash equilibrium. If the game does have a pure strategy Nash equilibrium, state the Nash equilibrium. Assume the row player is maximizing his playoffs which are showing in the matrices below.

\begin{table}[!h]
\centering
\begin{tabular}{lllc}
 &  & \multicolumn{2}{l}{Colin} \\ \cline{3-4}
 &  & C1 & \multicolumn{1}{l}{C2} \\ \hline
Rose & R1 & \multicolumn{1}{c}{10} & 10 \\
 & R2 & \multicolumn{1}{c}{5} & 0 \\ \hline
\end{tabular}
\end{table}


```{r, fig.height= 3, warning=FALSE, message=FALSE, cache=TRUE, tidy=TRUE}
library(DiagrammeR)
grViz("digraph boxes {

  graph [layout = neato, overlap = true, outputorder = edgefirst]

  node [shape = box]

  A [pos = '-1, 1!', label = 'C1, R1 \n (10)']; 
  B [pos = ' -1, -1!', label = 'C1, R2 \n (5)']; 
  C [pos = ' 1, 1!',label = 'C2, R1 \n (10)']; 
  D [pos = ' 1, -1!', label = 'C2, R2 \n (0)']; 

  # several 'edge' statements
  B->A C->A A->C D->B D->C 
}
")
```

We do have a pure strategy Nash equilibrium of 10 as our arrows all point to one value, when Rose plays strategy 1 and Colin plays either strategy 1 or 2. For graphing simplicity I set the strategy to Rose 1 and Colin 1.  

\newpage

# Page 385: problem 1 c

Using the definition provided for the movement diagram, determine whether the following zero-sum games have a pure strategy Nash equilibrium. If the game does have a pure strategy Nash equilibrium, state the Nash equilibrium. Assume the row player is maximizing his playoffs which are showing in the matrices below.

\begin{table}[!h]
\centering
\begin{tabular}{lllc}
 &  & \multicolumn{2}{l}{Pitcher} \\ \cline{3-4}
 &  & Fastball & \multicolumn{1}{l}{Knuckleball} \\ \hline
Batter & GFast & \multicolumn{1}{c}{.400} & .100 \\
 & Gknuckle & \multicolumn{1}{c}{.300} & .250 \\ \hline
\end{tabular}
\end{table}

```{r, fig.height= 3, tidy=TRUE, cache=TRUE, warning=FALSE, message=FALSE}
grViz("digraph boxes {

  graph [layout = neato, overlap = true, outputorder = edgefirst]

  node [shape = box]

  A [pos = '-1, 1!', label = 'Fastball, GFast \n (.400)']; 
  B [pos = ' 1, 1!', label = 'Knuckle, GFast  \n (.100)']; 
  C [pos = ' -1, -1!', label = 'Fast, GKnuckle \n (.300)']; 
  D [pos = ' 1, -1!', label = 'Knuckle, GKnuckle \n (.250)']; 

  # several 'edge' statements
  C->D A->B C->A B->D
}
")
```

The EV = .250 when the pitcher pitches a knuckle and the batter guesses a knuckle ball.  

\newpage

# Page 404: problem 2 a

For problems a-g build a linear programming model for each player's decisions and solve it both geometrically and algebraically. Assume the row player is maximizing his playoffs which are showing in the matrices below. 

\begin{table}[!h]
\centering
\begin{tabular}{lllc}
 &  & \multicolumn{2}{l}{Colin} \\ \cline{3-4}
 &  & C1 & \multicolumn{1}{l}{C2} \\ \hline
Rose & R1 & \multicolumn{1}{c}{10} & 10 \\
 & R2 & \multicolumn{1}{c}{5} & 0 \\ \hline
\end{tabular}
\end{table}

Geometrically: 

R1 has probability $p = x$, therefore the probability of R2 is $p=1-x$. 

$V_{C1}$ $\leq$ 10x + 5(1-x) = 5x + 5      
$V_{C2}$ $\leq$ 10x + 0(1-x)= 10x       
$x \leq 1$      
$x \geq 0$     

```{r}
library(knitr)
x <- c(0, 1, 0, 1)
A <- c(5, 10, 0, 10)
kable(as.data.frame(cbind(x,A)))
```


```{r, fig.height=2, fig.width=6, cache=TRUE, warning=FALSE, message=FALSE}
library(ggplot2)
plotc <- ggplot(data.frame(x=c(0,1)), aes(x)) +
          stat_function(fun=function(x)5*x+5, geom="line", aes(colour = "C1")) +
          stat_function(fun=function(x)10*x, geom="line", aes(colour = "C2")) +
            geom_vline(xintercept = 1) + 
          geom_vline(xintercept = 0)
plotc
```

Both are maxed at x = 1, $V_{C1}$ = 10, $V_{C2}$ = 10.

\newpage

R1 has probability $p = x$, therefore the probability of R2 is $p=1-x$. 

$V_{R1}$ $\leq$ 10x + 10(1-x) = 10
$V_{R2}$ $\leq$ 5x + 0(1-x)= 5x
$x \leq 1$
$x \geq 0$

```{r}
library(knitr)
x <- c(0, 1, 0, 1)
A <- c(10, 10, 0, 5)
kable(as.data.frame(cbind(x,A)))
```


```{r,  cache=TRUE, message=TRUE, warning=TRUE, fig.height=2, fig.width=6}

plotr <- ggplot(data.frame(x=c(0,1)), aes(x)) +
          stat_function(fun=function(x)5*x, geom="line", aes(colour = "R1")) +
          stat_function(fun=function(x)10, geom="line", aes(colour = "R2")) + 
            geom_vline(xintercept = 1) + 
          geom_vline(xintercept = 0)
plotr
```
\newpage

Now we plot both strategies for Rose and Colin. 

```{r,  cache=TRUE, message = TRUE, warning=TRUE , fig.height=3}
library(ggplot2)
plotrc <- ggplot(data.frame(x=c(0,1)), aes(x)) +
          stat_function(fun=function(x)5*x, geom="line", aes(colour = "R1")) +
          stat_function(fun=function(x)10, geom="line", aes(colour = "R2")) +
          stat_function(fun=function(x)5*x + 5, geom="line", aes(colour = "C1")) +
          stat_function(fun=function(x)10*x, geom="line", aes(colour = "C2")) +
          geom_vline(xintercept = 1) + 
          geom_vline(xintercept = 0)
plotrc
```

$V_{R1}$ is a constant at 10, whereas, $V_{R2}$ is maxed at x = 1; $V_{R2}$ = 5. 

We can therefore determine that Rose will play strategy R1 and Colin can play either strategy C1 or C2. 


# Page 420: problem 1

In the following problems, use the maximim and minimax method and movement diagram to determine if any pure strategy solution exist. Assume the row player is maximizing his payoffs which are shown in the matrices below. 

\begin{table}[h]
    \begin{tabular}{lllcc}
    ~    & ~  & Colin & ~  & Row Minimum  \\
    ~    & ~  & C1    & C2 & ~            \\
    Rose & R1 & 10    & 10 & \textbf{10}   \\
    ~    & R2 & 5     & 0  & \textbf{0}   \\ \hline
    \end{tabular}
\end{table}

\begin{table}[h]
    \begin{tabular}{lllc}
    ~    & ~  & Colin & ~   \\
    ~    & ~  & C1    & C2  \\
    Rose & R1 & 10    & 10  \\
    ~    & R2 & 5     & 0     \\ 
    Column Maximum  & & \textbf{10} & \textbf{0}
    \\ \hline
    \end{tabular}
\end{table}

As we can see from the row minimum that Colin can play either C1 or C2; and the column maximum indicates that Rose plays R1 which would be a pure strategy. 


