---
title: "9.4 Discussion"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(scales)
```

9.4 Implement a random walk Metropolis sampler for generating the standard Laplace distribution (see Exercise 3.2). For the increment, simulate from a normal distribution. Compare the chains generated when different variances are used for the proposal distribution. Also, compute the acceptance rates of each chain.

The standard Laplace distribution which has density $f(x) = \frac{1}{2}e^{-|x|},x \in I\!R$


```{r}
laplace_dist <- function(x){return((1/2)*exp(-abs(x)))}

RW.Metro.Laplace <- function(sigma, x.0, N) { 
                      x <- numeric(N) 
                      x[1] <- x.0 
                      u <- runif(N) 
                      k <- 0 
                      for (i in 2:N) { 
                        y <- rnorm(1, x[i-1], sigma) #increment
                        if (u[i] <= (laplace_dist(y) / laplace_dist(x[i-1]))) #proposal 
                            x[i] <- y 
                        else { 
                          x[i] <- x[i-1] 
                             k <- k + 1 } 
                      } 
                      return(list(x = x, k = k)) 
                    }
```

```{r}
N <- 200
sigma <- c(1/3, 2/3, 1, 3)
x.0 <- 100
results <- NULL 

for (i in 1:length(sigma)){
  a <- c(.05, seq(.1, .9, .1), .95)
  results[[i]] <- c(sigma_used = sigma[i], N = N, result = 
                    list(RW.Metro.Laplace(sigma = sigma[i], x.0 = x.0, N = N))) %>% 
                  c(., accepted.perc = scales::percent(1-unlist(.[[3]]['k'])/N),
                       quantile = list(quantile(unlist(.[[3]]['x']), a)))
}
```

```{r}
results %>% map(., function(x) c(x[1], x[4], x[5]))
```

Rizzo; Maria L.. Statistical Computing with R (Chapman & Hall/CRC The R Series) (Page 254). Taylor and Francis CRC ebook account. Kindle Edition. 