---
title: "Project 3 - Section 1"
author: "Team 1"
date: "March 23, 2016"
output:
  html_document:
    css: ./project.css
    highlight: pygments
    theme: cerulean
---

##4. Visualization and Analysis

###a. Visualization and Graphs

```{r, pre-load, include = FALSE}
library(RMySQL)
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(knitr)
library(wordcloud)

proj_user <- "project3"
proj_pwd  <- "CUNYRBridge4"
proj_db   <- "skill"
proj_host <- "db4free.net"

```

We used the Data Science Skills Database to create bar charts, word clouds and other visualization tools that show and summarize the group's findings. The Transformer group provided database `views` to provide  results to reduce the need for transformations in R. 

```{r}
skilldb <- dbConnect(MySQL(), user=proj_user, password=proj_pwd, dbname=proj_db, host=proj_host)
Views <- dbGetQuery( skilldb,"SELECT TABLE_NAME 'Views' 
                              FROM INFORMATION_SCHEMA.TABLES 
                              WHERE table_schema = 'skill' and table_type = 'view';")
kable(Views)
```

*** 

The Transformer team provided the skill names using two different methods. One involved a `scalar rating` that normalizes skill names across the entire database and the other used a `weighted rating` that weighted the ratings within their respective groups. 

We found that the top Data Science Skills are different for the 3 sources. We took the top three skills for each source from the `r Views[5,]` table. 

```{r}
Top3_by_source <- dbGetQuery( skilldb, sprintf("SELECT * FROM %s", Views[5,]))

Top3_by_source <- Top3_by_source %>%
                  arrange(source_name, desc(rating_scalar)) %>%
                  group_by(`source_name`) %>%
                  top_n(n = 3, wt = rating_scalar) %>%
                  select(source_name, skill_name, rating_scalar) 

kable(Top3_by_source)
```

Unsuprisingly, we see that technicial skills appear at the top of the lists for each of our data sources but not the exact same type.  

Let's look at the top 5 skills sets by skill type by pulling down the view from the database at `r  Views[9,]` and checking the results.

```{r}
scalar_skills <- dbGetQuery(skilldb, sprintf("Select * from %s", Views[9,]))

scalar_skills <- scalar_skills %>%
                   select(skill_type_name, rating_scalar, skill_set_name) %>%
                   arrange(skill_set_name)

ggplot(scalar_skills, aes(y = rating_scalar,
                          x = reorder(skill_set_name, -rating_scalar),  fill = skill_type_name)) + 
      geom_bar(stat = "identity")  + 
      scale_fill_discrete(name="Skill Type Name") + 
      xlab("Skill Set") + 
      ylab("Rating") + 
      theme(legend.position = "bottom",  
            axis.text.x = element_text(angle = 65, hjust = 1)) +
      ggtitle("Top 5 Skills by Skill Type")  
      
```

We continue to see techinical skills being the most often found skills in each source. `Big and Distributed Data` is the top skill amoung skill types which seems appropriate for data science.


####2. Horizontal Bar Graph Showing the Top 20 Data Science Skills from the 3 Sources (Google, Indeed and Kaggle)

Comparing the top lists of each data source will be useful in understanding some similarities or dissimalities between the sources

order by source_name (asc) and weighted_rating_overall
```{r}
jobdata <- dbGetQuery(skilldb, "select * from tbl_data")
newjobdata <- jobdata[with(jobdata, order(source_name,-weighted_rating_overall)),]
```

list skill_name for each source_name
```{r}
Google <- subset(newjobdata, source_name == "Google", select=c(source_name, skill_name,weighted_rating_overall))
Google <- Google[c(1:20),]
Indeed <- subset(newjobdata, source_name == "Indeed", select=c(source_name, skill_name, weighted_rating_overall))
Indeed <- Indeed[c(1:20),]
Kaggle <- subset(newjobdata, source_name == "Kaggle", select=c(source_name, skill_name, weighted_rating_overall))
Kaggle <- Kaggle[c(1:20),]
Combined <- cbind(Google,Indeed,Kaggle)
Combined$source_name <- NULL
Combined$source_name <- NULL
Combined$source_name <- NULL
colnames(Combined)[1] <- "GoogleSkills"
colnames(Combined)[2] <- "GoogleRatings"
colnames(Combined)[3] <- "IndeedSkills"
colnames(Combined)[4] <- "IndeedRatings"
colnames(Combined)[5] <- "KaggleSkills"
colnames(Combined)[6] <- "KaggleRatings"
kable(Combined)
```

plot results
```{r}
darkcols <- brewer.pal(8,"Dark2")
names <- Combined$GoogleSkills
barplot(Combined$GoogleRatings,main="GoogleRatings", horiz=TRUE, names.arg=names, las=1, col=darkcols, cex.axis=0.5, cex.names = 0.5)
names <- Combined$IndeedSkills
barplot(Combined$IndeedRatings,main="IndeedRatings", horiz=TRUE, names.arg=names, las=1, col=darkcols, cex.axis=0.5, cex.names = 0.5)
names <- Combined$KaggleSkills
barplot(Combined$KaggleRatings,main="KaggleRatings", horiz=TRUE, names.arg=names, las=1, col=darkcols, cex.axis=0.5, cex.names = 0.5)
```


####3. Bubble Graph Showing Weighted Rank of Skills by Skill Type and Source (Google, Indeed and Kaggle)

We have a wealth of information within our Data Science Skills database, we need to be able to quickly understand where the majority of highly ranked skills are located. The below graph provides a high level overview of the skills and how they relate to each other across the skill types by each source. 

```{r, fig.height=8, width = 11}
data_science <- jobdata

data_science$skill_name <- factor(data_science$skill_name)
data_science$skill_name <- factor(data_science$skill_name, levels = rev(levels(data_science$skill_name)))

ggplot(data_science[(data_science$source_id != 2),], 
       aes( source_name, skill_name, label = skill_name, 
           size = rating_scalar, fill = skill_type_name)) + 
        geom_point(pch = 21) +
        scale_fill_manual(values =  brewer.pal(9, "Set1")) + 
        scale_size_continuous(range =c(1,10)) + 
        facet_grid(~skill_type_name) + 
        theme_light() +
        xlab("Source") + 
        ylab("Skill") + 
        theme(legend.position = "none" , 
              axis.text.y = element_text(size=5), 
              axis.text.x = element_text(angle = 45, hjust = 1)) +
        ggtitle("Weighted Rank of Skill by Skill Type and Source")
```

It is clear that programming skills represent the majority of valuable skills that are highly sought after, followed by math and business. One limitation of this representation is while we can see where a majority of the highly ranked skills exist, we are not able to discern the highly ranked skills that are rare within the field. 


####4. Graph Showing Average Weighted Overall Rating by Source and Skill Type

```{r, eval=FALSE}
stab <- data %>% 
        group_by(source_name,skill_type_name) %>% 
        summarise(ave_wgt =mean(weighted_rating_overall))

ggplot(stab, aes(x =source_name, y=round(ave_wgt,2), fill = skill_type_name)) +
       geom_bar(stat="identity",position="dodge") + 
       xlab("Source") + 
       ylab("Average Weighted Rating Overall") + 
       ggtitle("Average Weighted Overall Rating by Source and Skill Type") 

```



```{r, message=FALSE, warning=FALSE}
require(wordcloud)

google <- subset(jobdata, source_name=="Google")
indeed <- subset(jobdata, source_name=="Indeed")
kaggle <- subset(jobdata, source_name=="Kaggle")

wordcloud(google$skill_name, google$weighted_rating_overall, random.order = FALSE, colors=brewer.pal(9, "YlGnBu"))
wordcloud(indeed$skill_name, log10(indeed$weighted_rating_overall), random.order = FALSE, colors=brewer.pal(8, "Dark2"))
wordcloud(kaggle$skill_name, kaggle$weighted_rating_overall, random.order = FALSE, colors=brewer.pal(8, "Dark2"))

```