---
title: "Homework 4"
author: "Christophe Hunt"
date: "February 20, 2017"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

\newpage

# Page 191: problem 3

Using Monte Carlo Simulation, write an algorithm to calculate an approximation to $\pi$ by considering the number of random points selected inside the quarter circle 
$$Q:x^2 +y^2 = 1, x \geq 0, y \geq 0$$
where the quarter circle is taken to be inside the square
$$S : 0 \leq x \leq 1~and~0 \leq y \leq 1$$
Use the equation $\frac{\pi}{4}=area$; $\frac{Q}{area}S$. 

```{r}
set.seed(1234)

monte_carlo <- function(n){
  counter = 0
  for (i in 1:n){
    y <- runif(1, 0, 1)
    x <- runif(1, 0, 1)
    if ((x^2 + y^2) < 1){
      counter <- counter + 1 
    } else {
      counter <- counter
    }
  }
  return(counter)
}
```

```{r}
n <- 500
(monte_carlo(n)/n)*4
```

```{r}
n <- 50000
(monte_carlo(n)/n)*4
```

\newpage

# Page 194: problem 1

Use the middle-square method to generate.

```{r}
middle_square <- function(n, seed) {
          suppressMessages(require(stringr))
          results <- list("instances" = n, "starting_seed" = seed)
          for (i in 1:n){
            j <- (seed^2)
              if (nchar(j) < 8){
                  j <- str_pad(j, 8, pad = "0")
              } else if (nchar(j) > 8) {
                j <- substr(j, 1, 8)
              }
            t <- j 
            j <- substr(j, 3, 6)
            seed <- as.numeric(j)
            results[[length(results)+1]] <- list("number" = t, "new_seed" = seed)
          }
         return(results)
}
```

## a.  10 random numbers using $x_0 = 1009$

```{r}
middle_square(10, 1009)
```

## b. 20 random numbers using $x_0 = 653217$

```{r, results='asis'}
middle_square(20, 653217)
```

## c. 15 random numbers using $x_0 = 3043$

```{r, results='asis'}
middle_square(15, 3043)
```
## TODO c. Comment about the results of each sequence. Was there cycling? Did each sequence degenerate rapidly? 


# Page 199: problem 4

Given loaded dice according to the following distribution, use Monte Carlo simulation to simulate the sum of 300 rolls of two unfair dice. 

```{r, results='asis'}
library(knitr)
Roll <- c(1:6)
Die_1 <- c(.1,.1,.2,.3,.2,.1)
Die_2 <- c(.3,.1,.2,.1,.05,.25)
kable(as.data.frame(cbind(Roll, Die_1, Die_2)))
```

```{r}
unfair_roll <- function(n, probs){
    if (length(probs) != 6){
      print("missing probabilities for all 6 sides")
      break
    }
    results <- list("1" = 0, "2" = 0 , "3" = 0, "4" = 0, "5" = 0, "6" = 0)
         for (i in 1:n){
       j <- 1 
       x <- runif(j, 0, j)
       if (x == 0){
        results[[j]] <- results[[j]] + 1
       } else if (x > 0 & x <= probs[[j]]){
        results[[j]] <- results[[j]] + 1
        next
       } else {
       j <- j + 1
       while ((x > sum(probs[0:(j-1)]) & x <= sum(probs[0:j])) != TRUE) {
       j <- j + 1
       }
       results[[j]] <- results[[j]] + 1
       }
     }
    for (i in 1:6){
     names(results[[i]]) <- "occurences"
     results[[i]][["probability"]] <- round(results[[i]]/n,2)
    }
  return(results)
}
```

```{r}
unfair_roll(300, Die_1)
```

```{r}
unfair_roll(300, Die_2)
```


# Page 211: problem 3

In many situations, the time $T$ between deliveries and the order quantity $Q$ is not fixed. Instead, an order is placed for a specific amount of gasoline. Depending on how many orders are placed in a given time interval, the time to fill an order varies. You have no reason to believe that the performance of the delivery operation will change. Therefore, you hvae examined records for the past 100 deliveries and found the following lag times, or extra days, required to fill your order:

```{r}
Lag_time <- c(2:7)
number_of_occurences <- c(10, 25, 30, 20, 13, 2)
X <- as.data.frame(cbind(Lag_time, number_of_occurences))
kable(rbind(X, c(" ", sum(number_of_occurences))))
```

Construct a Monte Carlo simulation for the lag time submodel. If you have a handheld calculator or computer avaialble, test your submodel by running 1000 trials and comparing the number of occurances of the various lag times with the historical data. 

```{r}
Lag_time <- c(2:7)
number_of_occurences <- c(10, 25, 30, 20, 13, 2)

lag_time <- function(n, probs, outcomes){
        results <- list()
        for (i in 1:length(outcomes)){
         results[[i]] <- 0
         names(results[[i]]) <- (paste("# of deliveries on day", outcomes[[i]]))
        }
     for (i in 1:n){
       j <- 1 
       x <- runif(j, 0, j)
       if (x == 0){
        results[[j]] <- results[[j]] + 1
       } else if (x > 0 & x <= probs[[j]]){
        results[[j]] <- results[[j]] + 1
        next
       } else {
       j <- j + 1
       while ((x > sum(probs[0:(j-1)]) & x <= sum(probs[0:j])) != TRUE) {
       j <- j + 1
       }
       results[[j]] <- results[[j]] + 1
       }
     }
    for (i in 1:length(results)){
     results[[i]][["probability"]] <- round(results[[i]]/n,2)
    }
  return(results)
}
probs <- (number_of_occurences/sum(number_of_occurences))
lag_time(1000, probs = probs , outcomes = Lag_time)
```
