---
title: "Homework 4"
author: "Christophe Hunt"
date: "February 20, 2017"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

\newpage

# Page 191: problem 3

Using Monte Carlo Simulation, write an algorithm to calculate an approximation to $\pi$ by considering the number of random points selected inside the quarter circle 
$$Q:x^2 +y^2 = 1, x \geq 0, y \geq 0$$
where the quarter circle is taken to be inside the square
$$S : 0 \leq x \leq 1~and~0 \leq y \leq 1$$
Use the equation $\frac{\pi}{4}=area$; $\frac{Q}{area}S$. 

```{r}
set.seed(1234)

monte_carlo <- function(n){
  counter = 0
  for (i in 1:n){
    y <- runif(1, 0, 1)
    x <- runif(1, 0, 1)
    if ((x^2 + y^2) < 1){
      counter <- counter + 1 
    } else {
      counter <- counter
    }
  }
  return(counter)
}
```

```{r}
n <- 500
(monte_carlo(n)/n)*4
```

```{r}
n <- 50000
(monte_carlo(n)/n)*4
```

\newpage

# Page 194: problem 1

Use the middle-square method to generate.

```{r}
middle_square <- function(n, seed) {
          suppressMessages(require(stringr))
          results <- list("instances" = n, "starting_seed" = seed)
          for (i in 1:n){
            j <- (seed^2)
              if (nchar(j) < 8){
                  j <- str_pad(j, 8, pad = "0")
              } else if (nchar(j) > 8) {
                j <- substr(j, 1, 8)
              }
            t <- j 
            j <- substr(j, 3, 6)
            seed <- as.numeric(j)
            results[[length(results)+1]] <- list("number" = t, "new_seed" = seed)
          }
         return(results)
}
```

## a.  10 random numbers using $x_0 = 1009$

```{r, results='asis'}
middle_square(10, 1009)
```

## b. 20 random numbers using $x_0 = 653217$

```{r, results='asis'}
middle_square(20, 653217)
```

## c. 15 random numbers using $x_0 = 3043$

```{r, results='asis'}
middle_square(15, 3043)
```
## TODO c. Comment about the results of each sequence. Was there cycling? Did each sequence degenerate rapidly? 


# Page 199: problem 4

Given loaded dice according to the following distribution, use Monte Carlo simulation to simulate the sum of 300 rolls of two unfair dice. 

```{r, results='asis'}
library(knitr)
Roll <- c(1:6)
Die_1 <- c(.1,.1,.2,.3,.2,.1)
Die_2 <- c(.3,.1,.2,.1,.05,.25)
kable(as.data.frame(cbind(Roll, Die_1, Die_2)))
```


# Page 211: problem 3



# Page 221: problem 2

