---
title: "Assignment 11"
author: "Christophe Hunt"
date: "April 23, 2017"
output: pdf_document
---

```{r, include = FALSE}
library(stargazer)
```

Using R's lm function, perform regression analysis and measure the signifiance of the independent variables for the following two data sets. 

In the first case, you are evaluating the statement that we hear that Maximum Heart Rate of a person is related to their age by the following equation:

MaxHR = 220 - Age  

You have been given the following sample:

Age 18 23 25 35 65 54 34 56 72 19 23 42 18 39 37     
MaxHR 202 186 187 180 156 169 174 172 153 199 193 174 198 183 178  

Perform a linear regression analysis fitting the Max Heart Rate to Age using the lm function in R. What is the resulting equation? 

```{r, results='asis'}
library(stargazer)
Age <- c(18, 23, 25, 35, 65, 54, 34, 56, 72, 19, 23, 42, 18, 39, 37)
MaxHR <- c(202, 186, 187, 180, 156, 169, 174, 172, 153, 199, 193, 174, 198, 183, 178)
stargazer(lm(MaxHR~Age), header = FALSE)
```

Is the effect of Age on Max HR significant? 

> Yes, the effect of Age on MaxH

What is the significance level? 

Please also plot the fitted relationship between Max HR and Age. 

Using the Auto data set from Assignment 5 (also attached here) perform a Linear Regression analysis using mpg as the dependent variable and the other 4 (displacement, horse-power, weight, acceleration) as independent variables. What is the final linear regression fit equation? 

Which of the 4 independent variables have a significant impact on mpg? What are their corresponding significance levels? 

What are the standard errors on each of the coefficients? 

Please perform this experiment in two ways. First take any random 40 data points from the entire auto data sample and perform the linear regression fit and measure the 95% confidence intervals. Then, take the entire data set (all 392 points) and perform linear regression and measure the 95% confidence intervals. Please report the resulting fit equation, their significance values and confidence intervals for each of the two runs. Please submit an R-markdown file documenting your experiments. Your submission should include the final linear fits, and their corresponding significance levels. In addition, you should clearly state what you concluded from looking at the fit and their significance levels.