---
title: "CHunt_Assignment2_PS1_PS2"
author: "Christophe Hunt"
date: "February 8, 2017"
output: pdf_document
---

## Problem Set 1 

1. Show that $A^TA\neq AA^T$ in general. (Proof and demonstration.)

For a 2x3 matrix A:
```{r}
A <- matrix(c(1,2,3,4,5,6), nrow=2, ncol=3)
A
```

The inverse of this matrix is:
```{r}
t(A)
```

The product of the inverse of the matrix and the matrix is :
```{r}
t(A)%*%A
```

Whereas, the product of the matrix and the inverse is : 

```{r}
A%*%t(A)
```

Therefore, $A^TA\neq AA^T$. 

\newpage

2. For a special type of square matrix A, we get $A^TA= AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

The conditions under which these are true are for special square matrix. In these instances, the inverse of the matrix is equal to the matrix $A^T = A$. Therefore, if $A^T = A$ then $A^TA= AA^T$. 

Such as the matrix :

```{r}
A <- matrix(c(1,2,3,2,1,2,3,2,1), nrow=3, ncol=3)
A
```

The inverse of this matrix is:
```{r}
t(A)
```

The product of the inverse of the matrix and the matrix is :
```{r}
t_AA <- t(A)%*%A
t_AA
```

The product of the matrix and the inverse is : 

```{r}
A_At <- A%*%t(A)
A_At
```

```{r}
A_At == t_AA
```


## Problem Set 2 

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. 

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. 

```{r}
factorization <- function(M){
  if(nrow(M) == 2){
     if(all(M[, 1] == 0)){
           return(print("cannot solve"))
          }
     while (M[1,1] == 0){
            M <- M[c(2,1), ]
          }
         E21 <- diag(nrow(M))
         E21[2,1] <- -M[2,1] / M[1,1]
         U <- E21 %*% M
         L <- solve(U)
         print(U)
         print(L)
      } else {
         if(all(M[, 1] == 0)){
           return(print("cannot solve"))
          }
         while (M[1,1] == 0){
            M <- M[c(2,3:nrow(M),1), ]
         }
         E21 <- diag(nrow(M))
         E21[2,1] <- -M[2,1] / M[1,1]
         M2 <- E21 %*% M
         E31 <- diag(nrow(M))
         E31[3,1] <- -M2[3,1] / M2[1,1]
         M3 <- E31 %*% M2
         E32 <- diag(nrow(M))
         E32[3,2] <- -M3[3,2] / M3[2,2]
         U <- E32 %*% M3
         L <- solve(E21) %*% solve(E31) %*% solve(E32)
         return(list("U" = U, "L" = L))
         
         }
      }
```



```{r}
A <- matrix(c(1,2,3,1,1,1,2,0,1),nrow=3)
factored <- factorization(A)
```

```{r}
factored$U
```

```{r}
factored$L
```
