---
title: "CHunt_Assignment2_PS1_PS2"
author: "Christophe Hunt"
date: "February 8, 2017"
output: pdf_document
---

## Problem Set 1 

1. Show that $A^TA\neq AA^T$ in general. (Proof and demonstration.)

If $A$ is a 3x2 matrix then $A^T$ is a 2x3 matrix. 

\[
\begin{bmatrix}
    x_{11} & x_{12} & \dots  & x_{1n} \\
    x_{21} & x_{22} & \dots  & x_{2n} \\
    \vdots & \vdots & \ddots & \vdots \\
    x_{d1} & x_{d2} & \dots  & x_{dn}
\end{bmatrix}
\]

For the matrix A below
```{r}
A <- matrix(c(1,2,3,4,5,6), nrow=2, ncol=3)
A
```

The inverse of this matrix is 
```{r}
t(A)
```

The product of the inverse of the matrix and the matrix is as per below.
```{r}
t(A)%*%A
```

Whereas, the product of the matrix and the inverse is as per below. 

```{r}
A%*%t(A)
```

Therefore, $A^TA\neq AA^T$. 

2. For a special type of square matrix A, we get $A^TA= AA^T$. Under what conditions could this be true? (Hint: The Identity matrix I is an example of such a matrix).

The conditions under which these are true are for special square matrix. In these instances, the inverse of the matrix is equal to the matrix $A^T = A$. Therefore, if $A^T = A$ then $A^TA= AA^T$. 


## Problem Set 2 

Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars. 

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer. Please submit your response in an R Markdown document using our class naming convention, E.g. LFulton_Assignment2_PS2.png You don't have to worry about permuting rows of A and you can assume that A is less than 5x5, if you need to hard-code any variables in your code. If you doing the entire assignment in R, then please submit only one markdown document for both the problems.

```{r}


```





