---
title: "Homework 3"
author: "Christophe Hunt"
date: "February 15, 2017"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

# Problem Set 1

## Problem 1 

What is the rank of the matrix A?

\[
A = \begin{bmatrix}
    1 & 2 & 3 &4  \\
   -1 & 0 & 1 & 3  \\
   0 & 1 & -2 & 1   \\
   5 & 4 & -2 & -3   \\
\end{bmatrix}
\]


```{r}
A <- t(matrix(c(1 , 2 ,3 , 4, 
              -1 , 0 , 1 , 3, 
              0 , 1 , -2 , 1, 
              5 , 4 , -2 , -3), nrow = 4, ncol = 4))

```
Since A is a square matrix of (`r paste(as.integer(ncol(A)), "x",  as.integer(nrow(A)))`) and the determinate is  `r det(A)` which is $\neq$ 0, the rank is simply `r as.integer(ncol(A))`

Below is my attempt to create a function to rank matrix A and matrix B in problem 3 programmatically. I feel that the part of the function that calculates the determinates of the submatrices could be improved. 

```{r}
A <- t(matrix(c(1 , 2 ,3 , 4, 
              -1 , 0 , 1 , 3, 
              0 , 1 , -2 , 1, 
              5 , 4 , -2 , -3), nrow = 4, ncol = 4))

subdet <- function(A, t){
            subdet <- 0
           for (i in 1:t){
             subdet <- rbind(subdet,det(A[-i,-i]))
           }
           if (sum(subdet) !=0){
             return(as.integer(ncol(A))-1)
           } else {
             0
           }
}

Matrix.Rank <- function(A){
               sq.matrix <- (as.integer(ncol(A)) == as.integer(nrow(A)))
               if (all(A == 0)){
                 return(0)
               } else if (sq.matrix == TRUE){
                 det.0 <- det(A) != 0
                 if (det.0 == TRUE) {
                 return(as.integer(ncol(A)))
                 } else if (subdet(A,ncol(A)) != 0) {
                  return(as.integer(ncol(A))-1)
                 } else {
                     return(1)
                   }
                 }
               }

Matrix.Rank(A)
```


## Problem 2 

Given an mxn matrix where $m > n$, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

> The maximum rank of a rectangular matrix is the maximum columns or rows for the lesser value. Therefore, given an mxn matrix where $m > n$, the maximum rank is $n$. 

> Assuming the rectangular matrix has at least one non-zero element, it's minimum rank must be greater than zero, therefore the minimum rank would be 1. 

## Problem 3 

What is the rank of matrix B?

\[
B = 
\begin{bmatrix}
   1 & 2 & 1 \\
   3 & 6 & 3 \\
   2 & 4 & 2 \\
\end{bmatrix}
\]

```{r}
B <- t(matrix(c( 1 , 2 , 1 ,  3 ,  6 , 3 ,  2 ,  4 , 2 ), ncol = 3, nrow = 3))
```

This matrix rows that are linearly dependent, R2 = `r B[2,]` and R3 = `r B[3,]` are mulitples of R1 = `r B[1,]`. This made my function more challenging because the determinates of the submatrices of B are also $= 0$. However, as long as there is at least one non-zero element in the matrix the minimum rank will be $= 1$.

```{r}
Matrix.Rank(B)
```

# Problem Set 2 

Compute the eigenvalues and eigenvectors of the matrix A. You'll need to show your work. You'll need to write out the characteristic polynomial and show your solution.

\[
A = 
\begin{bmatrix}
   1 & 2 & 3 \\
   0 & 4 & 5 \\
   0 & 0 & 6 \\
\end{bmatrix}
\]


\[
det( 
\begin{bmatrix}
   1 & 2 & 3 \\
   0 & 4 & 5 \\
   0 & 0 & 6 \\
\end{bmatrix}
-
\begin{bmatrix}
   \lambda & 0 & 0 \\
   0 & \lambda  & 0 \\
   0 & 0 & \lambda  \\
\end{bmatrix}
) = 0
\]


\[
det( 
\begin{bmatrix}
   1 - \lambda & 2 & 3 \\
   0 & 4 - \lambda & 5 \\
   0 & 0 & 6- \lambda\\
\end{bmatrix}
)
\]


```{r}
A <-  t(matrix(c(1 , 2 , 3,  0 , 4 , 5 , 0 , 0 , 6 ), nrow = 3, ncol=3))
```



