---
title: "Homework week 4"
author: "Christophe"
date: "February 15, 2016"
output: html_document
---
```{r preload library, include = FALSE}
library(stringr)
library(Hmisc)
```

```{r}
library(stringr)
library(Hmisc)
raw.data <- "555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5543642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))
last_first <- name[!is.na((str_extract(name, ".+, .+")))] #finds all names in "ln, fn"" format 
first_name <- (str_extract(last_first,", .+")) #gets just the first name from "ln, fn" format 
last_name <- (str_extract(last_first,".+, ")) #gets just the last name from "ln, fn" format 
names <- as.list(paste(first_name,last_name)) #puts the first name and last name together
last_first <- as.list(gsub(", ", "", names)) # removes left over "," in the names
first_last <- as.list(name[is.na((str_extract(name, ".+, .+")))]) # gets all names in "fn ln" format
name <- append(last_first, first_last) #combines the two lists
list.tree(name)
```

## b 
```{r}
title <- name[!is.na((str_extract(name, "[:alpha:]{2,}\\..")))] #gets all names that have more than 2 letters before period 
list.tree(title)
```

## c 

```{r}
three_names <- name[!is.na(str_extract(name, ("[:upper:]\\.+")))] # gets names with one capital letter before next name part
list.tree(three_names)
```

## 7 HTML tag problem

```{r}
string <- "<title>+++BREAKING NEWS+++</title>"
attempt_1 <- str_extract(string, "<.+>") 
attempt_1
attempt_2 <- str_extract(string, "<\\w+>")
attempt_2
```

The reason that `attempt_1` fails is that it is not attempting to grab a whole word within the first HTML tag. It is using a lazy grab and because there is another tag at the end of the HTML line it pulls the entire line (*I think*). 


## 8
```{r}
string <- "(5-3)^2=5^2-2*5*3+3^2"
attempt_1 <- str_extract(string, "[^0-9=+*()]+")
attempt_1
attempt_2 <- str_extract(string, "[0-9=+*()].+")
attempt_2
```

The reason that `attempt_1` fails is because the carat `^` symbol inverses the matching. So it is looking for anything no 0-9. Also the period `.` is any characters after the matching of one time using the `+` symbol. 

```{r}


```
