---
title: "Principles of Data Visualization & Intro to ggplot2"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
library(pacman)
p_load(plyr, dplyr, tidyr, ggplot2, rgdal, maptools, RColorBrewer, scales, tools)
```

```{r}
url_for_download <- c("https://raw.githubusercontent.com/jlaurito/CUNY_IS608/master/lecture1/data/inc5000_data.csv")
raw_data <- read.csv(url_for_download)
raw_data_complete_cases <- raw_data[(complete.cases(raw_data)),]
raw_data_name <- file_path_sans_ext(gsub(".+/", "", url_for_download))
business_data <- as.data.frame(raw_data_complete_cases %>%
                               select(Name, State) %>%
                               group_by(State) %>%
                               summarise(Total = length(Name))
                              )
```


```{r, results = "hide" }
state_shape_files <- "http://www2.census.gov/geo/tiger/GENZ2015/shp/cb_2015_us_state_20m.zip"
download.file(state_shape_files, destfile = "cb_2015_us_state_20m.zip")
unzip("cb_2015_us_state_20m.zip")
states <- readOGR(dsn = ".", layer = "cb_2015_us_state_20m")
```

I ran into alot of trouble with Alaska making the graph too wide so I used the below code snippet to move around Hawaii and Alaska which came from this stackoverflow post - http://stackoverflow.com/questions/13757771/relocating-alaska-and-hawaii-on-thematic-map-of-the-usa-with-ggplot2
```{r}
require(maptools)
require(rgdal)

fixup <- function(usa,alaskaFix,hawaiiFix){

  alaska=usa[usa$NAME=="Alaska",]
  alaska = fix1(alaska,alaskaFix)
  proj4string(alaska) <- proj4string(usa)

  hawaii = usa[usa$NAME=="Hawaii",]
  hawaii = fix1(hawaii,hawaiiFix)
  proj4string(hawaii) <- proj4string(usa)

  usa = usa[! usa$NAME %in% c("Alaska","Hawaii"),]
  usa = rbind(usa,alaska,hawaii)

  return(usa)

}

fix1 <- function(object,params){
  r=params[1];scale=params[2];shift=params[3:4]
  object = elide(object,rotate=r)
  size = max(apply(bbox(object),1,diff))/scale
  object = elide(object,scale=size)
  object = elide(object,shift=shift)
  object
}

states <- spTransform(states,CRS("+init=epsg:2163")) %>%
          fixup(c(-35,1.5,-2800000,-2600000),c(-35,1,6800000,-1600000))

```
end of code snippet from stack overflow 

```{r, results='hide'}
states@data$id <- rownames(states@data)

states.df <- fortify(states, region = "id") %>% 
             join(states@data, by = "id") %>%
             merge(business_data, by.x = c("STUSPS"), by.y = c("State"))

file.remove(dir(getwd(), pattern = "cb_2015_us_state_20m", full.names = TRUE)) #Clean Up 

bcounts_wNames <- fortify(aggregate(cbind(long, lat) ~ Total + NAME, data = states.df, FUN = mean)) 
```


```{r, fig.height= 7.5, fig.width=10}
ggplot(states.df) + 
  geom_polygon(colour="grey40", aes(long, lat, group = group, fill = Total) ) +
  geom_text(data = bcounts_wNames, aes(long, lat, label = Total), size=3, colour = "firebrick1", fontface = "bold") + 
  theme_void() +
  scale_fill_gradientn(colors =  brewer.pal(9,"PuBu"), space = "Lab", guide = "legend") +
  ggtitle(paste0("Count of Companies in Data Set : ", raw_data_name," by US State (including US Territory of Puerto Rico) \n"))
```

## Let's dig in on the State with the 3rd most companies in the data set. Imagine you work for the state and are interested in how many people are employed by companies in different industries employ. Create a plot of average employment by industry for companies in this state (only use cases with full data (user R's complete.cases() function). Your graph should show how variable the ranges are, and exclude outliers.


```{r}
library(ggplot2)

state_most_companies <- 3

third_most_companies <- as.data.frame(raw_data_complete_cases %>%
                                         select(Name, State) %>%
                                         group_by(State) %>%
                                         summarise(Total = length(Name)) %>%
                                         arrange(-Total)
                                        )[state_most_companies,]

business_data <- raw_data_complete_cases[(raw_data_complete_cases$State == third_most_companies$State),] %>%
                 select(Industry, Employees) %>%
                 group_by(Industry) %>%
                 mutate(no_of_companies = length(Industry)) %>%
                 mutate(is_outlier = ifelse(no_of_companies == 1 , 0, 
                                     ifelse(Employees > sd(Employees, na.rm = TRUE) * 3, 1, 0))) %>%
                 filter(is_outlier == 0) %>% 
                 mutate(avg_emp = mean(Employees), 
                        sem = ifelse(no_of_companies == 1 , 0, 
                                     sd(Employees, na.rm = TRUE)/sqrt(length(Employees))))  %>%
                 select(-Employees) %>%
                 distinct(avg_emp, sem)

ggplot(business_data, aes(x = Industry, y = avg_emp)) + 
       geom_bar(stat="identity") + 
       geom_errorbar(aes(ymin= avg_emp - sem, ymax=avg_emp + sem), width = .5) +  
       theme_minimal() + 
       theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
       ylab(label = "Average Number of Employees \n") +
       xlab(label = "\n Industry Type \n") +
       ggtitle(label = paste0("Average Number of Employees by Industry for ", third_most_companies$State))
```

3.	Now imagine you work for an investor and want to see which industries generate the most revenue per employee. Create a chart makes this information clear.

```{r}

revenue_data <- as.data.frame(raw_data_complete_cases %>%
                              select(Revenue, Employees, Industry) %>%
                              group_by(Industry) %>% 
                              mutate(Revenue_Per_Employee = (Revenue/Employees)) %>%
                              mutate(avg = mean(Revenue_Per_Employee),
                                     median = median(Revenue_Per_Employee)) %>%
                              gather(type, value, avg:median) %>%
                              select(Industry, type, value) %>%
                              distinct(type, value)
                              )

ggplot(revenue_data, aes(x = Industry, y = value, fill = type, color = type)) + 
       geom_point() + 
       theme_minimal() + 
       theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
       scale_color_discrete(name = "Employee Revenue Statistic") + 
       ylab(label = "Revenue\n") +
       xlab(label = "\n Industry Type") +
       scale_y_continuous(labels = dollar) + 
       ggtitle("Revenue Per Employee by Industry Type\n  \n")
```
