---
title: "Project 3 - Section 1"
author: "Team 1"
date: "March 23, 2016"
output:
  html_document:
    css: ./project.css
    highlight: pygments
    theme: cerulean
---

##4. Visualization and Analysis

###a. Visualization and Graphs

The Presenters used the  Data Science Skills Database to create bar charts, word clouds and other visualization tools that show and summarize the group's findings.

When we examined the data, we found the the skill names are weighted on different scales and that the top Data Science Skills are different for the 3 sources. 

```{r, include = FALSE}
library(RMySQL)
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(knitr)

proj_user <- "project3"
proj_pwd  <- "CUNYRBridge4"
proj_db   <- "skill"
proj_host <- "db4free.net"
```

```{r}
skilldb <- dbConnect(MySQL(), user=proj_user, password=proj_pwd, dbname=proj_db, host=proj_host)
weightedskills <- dbGetQuery(skilldb, "select * from tbl_data")
```

include all R packages needed for visualization of results
```{r}
library(tidyr)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(knitr)
```


convert columns to proper data type, pick relevant columns and arrange by source
```{r}
weightedskills$skill_name <- as.character(weightedskills$skill_name)
weightedskills$weighted_rating_overall <- as.numeric(weightedskills$weighted_rating_overall)
weightedskills$source_name <- as.character(weightedskills$source_name)

weightedskills1 <- weightedskills %>%
                   select(skill_name, weighted_rating_overall, source_name) %>%
                   arrange(source_name)
```

convert vector to a data frame
```{r}
weightedskills1 <- data.frame(weightedskills1)
```

pick all rows that have a Google source
```{r}
weightedskills11 <- filter(weightedskills1, source_name == "Google")
weightedskills11 <- weightedskills11[order(-weightedskills11$weighted_rating_overall),]
```

pick the top six Google data science skills and generate bar graph
```{r}
h1 = head(weightedskills11)
p1 = ggplot(h1, aes(y = weighted_rating_overall, fill = skill_name))
h1$skill_name = reorder(h1$skill_name, -h1$weighted_rating_overall)
p1 + geom_bar(aes(x = skill_name), data = h1, stat = "identity") + xlab("skill name") + ylab("weighted rating overall") + theme(legend.position = "none") +
ggtitle("Top 6 Data Science Skills - Google")
```

pick all rows that have a Indeed source
```{r}
weightedskills12 = filter(weightedskills1, source_name == "Indeed")
weightedskills12 = weightedskills12[order(-weightedskills12$weighted_rating_overall),]
```

pick the top six Indeed data science skills and generate bar graph
```{r}
h2 = head(weightedskills12)
p2 = ggplot(h2, aes(y = weighted_rating_overall, fill = skill_name))
h2$skill_name = reorder(h2$skill_name, -h2$weighted_rating_overall)
p2 + geom_bar(aes(x = skill_name), data = h2, stat = "identity") + xlab("skill name") + ylab("weighted rating overall") + theme(legend.position = "none") +
ggtitle("Top 6 Data Science Skills - Indeed")
```

pick all rows that have a Kaggle source
```{r}
weightedskills13 = filter(weightedskills1, source_name == "Kaggle")
weightedskills13 = weightedskills13[order(-weightedskills13$weighted_rating_overall),]
```

pick the top six Kaggle data science skills and generate bar graph
```{r}
h3 = head(weightedskills13)
p3 = ggplot(h3, aes(y = weighted_rating_overall, fill = skill_name))
h3$skill_name = reorder(h3$skill_name, -h3$weighted_rating_overall)
p3 + geom_bar(aes(x = skill_name), data = h3, stat = "identity") + xlab("skill name") + ylab("weighted rating overall") + theme(legend.position = "none") +
ggtitle("Top 6 Data Science Skills - Kaggle")
```



####2. Horizontal Bar Graph Showing the Top 20 Data Science Skills from the 3 Sources (Google, Indeed and Kaggle)


read data
```{r}
jobdata <- read.csv("https://raw.githubusercontent.com/danielhong98/MSDA-Spring-2016/347a383eae3b9f02bc5d128efb5de14e1f688f8e/tbl_data_v2.csv")
```

order by source_name (asc) and weighted_rating_overall
```{r}
newjobdata <- jobdata[with(jobdata, order(source_name,-weighted_rating_overall)),]
```

list skill_name for each source_name
```{r}
Google <- subset(newjobdata, source_name == "Google", select=c(source_name, skill_name,weighted_rating_overall))
Google <- Google[c(1:20),]
Indeed <- subset(newjobdata, source_name == "Indeed", select=c(source_name, skill_name, weighted_rating_overall))
Indeed <- Indeed[c(1:20),]
Kaggle <- subset(newjobdata, source_name == "Kaggle", select=c(source_name, skill_name, weighted_rating_overall))
Kaggle <- Kaggle[c(1:20),]
Combined <- cbind(Google,Indeed,Kaggle)
Combined$source_name <- NULL
Combined$source_name <- NULL
Combined$source_name <- NULL
colnames(Combined)[1] <- "GoogleSkills"
colnames(Combined)[2] <- "GoogleRatings"
colnames(Combined)[3] <- "IndeedSkills"
colnames(Combined)[4] <- "IndeedRatings"
colnames(Combined)[5] <- "KaggleSkills"
colnames(Combined)[6] <- "KaggleRatings"
kable(Combined)
```

plot results
```{r}
darkcols <- brewer.pal(8,"Dark2")
names <- Combined$GoogleSkills
barplot(Combined$GoogleRatings,main="GoogleRatings", horiz=TRUE, names.arg=names, las=1, col=darkcols, cex.axis=0.5, cex.names = 0.5)
names <- Combined$IndeedSkills
barplot(Combined$IndeedRatings,main="IndeedRatings", horiz=TRUE, names.arg=names, las=1, col=darkcols, cex.axis=0.5, cex.names = 0.5)
names <- Combined$KaggleSkills
barplot(Combined$KaggleRatings,main="KaggleRatings", horiz=TRUE, names.arg=names, las=1, col=darkcols, cex.axis=0.5, cex.names = 0.5)
```


####3. Bubble Graph Showing Weighted Rank of Skills by Skill Type and Source (Google, Indeed and Kaggle)

```{r}
data<- read.csv("https://raw.githubusercontent.com/ChristopheHunt/MSDA---Coursework/master/Data%20607/Homework/Group%20Project/tbl_data_version1%20.csv")
```

```{r}
ggplot(data, aes(source_name, skill_name, label = skill_name, 
                    size = weighted_rating_overall, fill = skill_type_name)) + 
        geom_point(pch = 21) + 
        scale_fill_manual(values =  brewer.pal(9, "Set1")) + 
        scale_size_continuous(range =c(1,20)) + 
        facet_grid(~skill_type_name) + 
        theme_light() +
        xlab("Source") + 
        ylab("Skill") + 
        theme(legend.position = "none" , axis.text.y = element_text(size=3)) +
        ggtitle("Weighted Rank of Skill by Skill Type and Source")
```

####4. Bubble Graph Showing Average Weighted Overall Rating by Source and Skill Type

```{r}
stab <- data %>% group_by(source_name,skill_type_name) %>% summarise(ave_wgt =mean(weighted_rating_overall))

ggplot(stab, aes(x =source_name, y=round(ave_wgt,2), fill = skill_type_name)) +  geom_bar(stat="identity",position="dodge") + xlab("Source") + ylab("Average Weighted Rating Overall") + ggtitle("Average Weighted Overall Rating by Source and Skill Type") 

```


Jeff?
Musa - ?
###a. Visualization and Graphs

###b. Conclusion






