---
title: "Exponential Smoothing"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r, include=FALSE}
library(fma)
library(tidyverse)
library(ggfortify)
```


# 7.1

Data set books contains the daily sales of paperback and hardcover books at the same store. The task is to forecast the next four days' sales for paperback and hardcover books (data set books).

a. Plot the series and discuss the main features of the data.

```{r, fig.height=4}
library(fma)
library(tidyverse)
library(ggfortify)
books <- fma::books
plot(books)
```

```{r}
autoplot(books)
```

> The data appears to have some seasonal variance and there appears to be times where Paperback is at the opposite extreme of the Hardcover books. 

b. Use simple exponential smoothing with the ses function (setting initial="simple") and explore different values of $\alpha$ for the paperback series. Record the within-sample SSE for the one-step forecasts. Plot SSE against $\alpha$  and find which value of $\alpha$  works best. What is the effect of $\alpha$  on the forecasts?

```{r}
paperback <- books[,1]

fit1 <- ses(paperback, alpha=0.4, initial="simple", h=3)
fit2 <- ses(paperback, alpha=0.6, initial="simple", h=3)
fit3 <- ses(paperback, alpha=0.8, h=3)

plot(fit1, ylab="Paperback Books Sold",
  xlab="Time Series", main="", fcol="white", type="o")
lines(fitted(fit1), col="blue", type="o")
lines(fitted(fit2), col="red", type="o")
lines(fitted(fit3), col="green", type="o")
lines(fit1$mean, col="blue", type="o")
lines(fit2$mean, col="red", type="o")
lines(fit3$mean, col="green", type="o")
legend("topleft",lty=1, col=c(1,"blue","red","green"), 
  c("data", expression(alpha == 0.4), expression(alpha == 0.6),
  expression(alpha == 0.8)),pch=1)
```

> Due to so much variability in the data, it's challenging to tell which worked best. The best performance would dependent on the use of the results, I would say $\alpha = 0.4$ looks to be the best predictor but $\alpha = 0.8$ appears to be a more conservative assumption for the next best value. 

c. Now let ses select the optimal value of $\alpha$  Use this value to generate forecasts for the next four days. Compare your results with 2.

```{r}
fit_opt <- ses(paperback, initial="simple", h=4)
plot(fit_opt, ylab="Paperback Books Sold",
  xlab="Time Series", main="", fcol="white", type="o")
lines(fitted(fit_opt), col="blue", type="o")
lines(fit_opt$mean, col="blue", type="o")
```

```{r}
summary(fit_opt)
```

> The optimal alpha for the paperback books is $\alpha$ = 0.2125 

d. Repeat but with initial="optimal". How much difference does an optimal initial level make?

```{r}
fit_opt <- ses(paperback, initial="optimal", h=4)
plot(fit_opt, ylab="Paperback Books Sold",
  xlab="Time Series", main="", fcol="white", type="o")
lines(fitted(fit_opt), col="blue", type="o")
lines(fit_opt$mean, col="blue", type="o")
```

```{r}
summary(fit_opt)
```

> The optimal alpha with the setting of optimal in ses = "optimal" is $\alpha = 0.1685$

e. Repeat steps (b)-(d) with the hardcover series.

### Hardcover books

b2. Use simple exponential smoothing with the ses function (setting initial="simple") and explore different values of $\alpha$ for the paperback series. Record the within-sample SSE for the one-step forecasts. Plot SSE against $\alpha$  and find which value of $\alpha$  works best. What is the effect of $\alpha$  on the forecasts?

```{r}
hardcover <- books[,2]

fit1 <- ses(hardcover, alpha=0.4, initial="simple", h=3)
fit2 <- ses(hardcover, alpha=0.6, initial="simple", h=3)
fit3 <- ses(hardcover, alpha=0.8, h=3)

plot(fit1, ylab="Hardcover Books Sold",
  xlab="Time Series", main="", fcol="white", type="o")
lines(fitted(fit1), col="blue", type="o")
lines(fitted(fit2), col="red", type="o")
lines(fitted(fit3), col="green", type="o")
lines(fit1$mean, col="blue", type="o")
lines(fit2$mean, col="red", type="o")
lines(fit3$mean, col="green", type="o")
legend("topleft",lty=1, col=c(1,"blue","red","green"), 
  c("data", expression(alpha == 0.4), expression(alpha == 0.6),
  expression(alpha == 0.8)),pch=1)
```

> Due to so much variability in the data, it's challenging to tell which worked best in the hardcover as well. The best performance would be what I was trying to achieve with my results, so I would say $\alpha = 0.4$ appears to be the safest assumption for the next best value but $\alpha = 0.8$ looks like it could be a more accurate prediction. 

c2. Now let ses select the optimal value of $\alpha$  Use this value to generate forecasts for the next four days. Compare your results with 2.

```{r}
fit_opt <- ses(hardcover, initial="simple", h=4)
plot(fit_opt, ylab="Hardcover Books Sold",
  xlab="Time Series", main="", fcol="white", type="o")
lines(fitted(fit_opt), col="blue", type="o")
lines(fit_opt$mean, col="blue", type="o")
```

```{r}
summary(fit_opt)
```

> The optimal alpha for the hardcover books is $\alpha$ = 0.3473$ which fits in line with my previous thoughts on best alpha

d. Repeat but with initial="optimal". How much difference does an optimal initial level make?

```{r}
fit_opt <- ses(hardcover, initial="optimal", h=4)
plot(fit_opt, ylab="Paperback Books Sold",
  xlab="Time Series", main="", fcol="white", type="o")
lines(fitted(fit_opt), col="blue", type="o")
lines(fit_opt$mean, col="blue", type="o")
```

```{r}
summary(fit_opt)
```

> The optimal alpha with the setting of optimal in ses = "optimal" is $\alpha = 0.3283$ which is still very close the assumptions made when $\alpha$ was set to 0.3283.

# 7.3

For this exercise, use the quarterly UK passenger vehicle production data from 1977:1--2005:1 (data set ukcars).

a. Plot the data and describe the main features of the series.

```{r}
library(expsmooth)
ukcars <- expsmooth::ukcars
```

b. Decompose the series using STL and obtain the seasonally adjusted data.

```{r}
fit_stl <- stl(ukcars, s.window = "periodic")
plot(fit_stl)
```

> We see a long gradual upwards trend from ~1980 to ~2000 followed by a drop and gradual trend upwards. 

```{r}
sadj <- seasadj(fit_stl)
sfactors <- fit_stl$time.series[2:11, "seasonal"] 
```


c. Forecast the next two years of the series using an additive damped trend method applied to the seasonally adjusted data. Then reseasonalize the forecasts. Record the parameters of the method and report the RMSE of the one-step forecasts from your method.

```{r}
fit_dpd_sadj <- holt(sadj, damped = TRUE)
paste0("Additive damped RMSE : ",round(accuracy(fit_dpd_sadj)[2], 3))
```

```{r}
rsfit_dpd_sadj <- fit_dpd_sadj$mean + sfactors 
rsfit_dpd_sadj
```

```{r}
plot(ukcars, 
     type = "o",
     xlab = "Years", 
     ylab = "UK Cars", 
     xlim = c(1997, 2008))

lines(rsfit_dpd_sadj, type = "o", col = "blue")
```

> The results are very impressive, it produced a very reasonable forcast excluding any shock events, which would be impossible to predict. 

d. Forecast the next two years of the series using Holt's linear method applied to the seasonally adjusted data. Then reseasonalize the forecasts. Record the parameters of the method and report the RMSE of of the one-step forecasts from your method.

```{r}
fit_hl <- holt(sadj)
paste0("Holt Linear Method with seasonal adjustment RMSE: ",round(accuracy(fit_hl)[2],3))
```

```{r}
autoplot(fit_hl, xlab = "Years", ylab = "UK Car Production")
```

> RSME 

```{r}
accuracy(fit_hl)
```

> Reseaonalize the forecast

```{r}
rs_l <- fit_hl$mean + sfactors
plot(ukcars, ylab = "UK Care Production")
lines(rs_l, type = "o", col = "red")
```


e. Now use ets() to choose a seasonal model for the data.

```{r}
fit_ets <- ets(ukcars, model = "ANN")
summary(fit_ets)
```

```{r}
autoplot(forecast(fit_ets), xlab = "Years", ylab = "UK Car Production")
```

```{r}
paste0("RMSE from Simple Exp Smoothing Method: ",round(accuracy(fit_ets)[2], 3))
```

f. Compare the RMSE of the fitted model with the RMSE of the model you obtained using an STL decomposition with Holt's method. Which gives the better in-sample fits?

```{r}
paste0("Additive Damped - RMSE : ",round(accuracy(fit_dpd_sadj)[2], 3))
paste0("Holt -  RMSE: ",round(accuracy(fit_hl)[2],3))
paste0("RMSE from Simple Exp Smoothing Method: ",round(accuracy(fit_ets)[2], 3))
```

g. Compare the forecasts from the three approaches? Which seems most reasonable?

```{r}
autoplot(forecast(fit_dpd_sadj))
autoplot(forecast(fit_hl))
autoplot(forecast(fit_ets))
```

> I find this a difficult evaluation on the most appropriate model. Using our evaluation methods the lowest RMSE is the Additive Damped model but visually the second Holt model appears to have the more accurate prediction. I think this is an interesting dilemma because the comparison measure indicates that the first model is best but the visuals of the second model look more sound and there is a conflict with my intiuition and the data. The first model with the lowest RMSE is the most reasonable choice. I did not discuss the simple exponential smoothing method because its performance is clearly subpar.
