---
title: "Week 8 Discussion Question"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

## Stratified Sampling 

The purpose behind stratified sampling is to split up our domain into regions and sample points from each region. Allowing each region to provide a proportional share of points and combining to get a better estimate of our example of $\int_{0}^{1} e^{-x}(1+x^2)^{-1} dx$. - http://statweb.stanford.edu/~owen/mc/Ch-var-basic.pdf

"We can achieve more reduction in variance using stratification when the means of the strata are widely disperses than if the means of the strata are approximately equal." - Rizzo, Maria L. Statistical computing with R. Chapman & Hall/CRC, 2008. Below is an example from the text. 

```{r}
set.strat <- list(M = 120, rows = 20)
T2 <- numeric(8)
estimates <- matrix(0, set.strat$rows, 2)
g <- function (x) {
       exp(-x - log(1+x^2)) * (x > 0) * (x <1)}

for(i in 1:set.strat$rows){
  estimates[i, 1] <- mean(g(runif(set.strat$M)))
  bin <- seq(0, 1, by = 1/length(T2))
  for (j in 1:length(T2)){
  T2[[j]] <- mean(g(runif(set.strat$M/length(T2), bin[[j]], bin[[j+1]])))
  }
  estimates[i,2] <- mean(T2)
}
```

```{r}
estimates
apply(estimates, 2, mean)
```

```{r}
options(scipen = 999)
apply(estimates, 2, var)
```

The text - http://statweb.stanford.edu/~owen/mc/Ch-var-basic.pdf - has a great example of stratified sample for Compound Poisson models for rainfall. 