---
title: "XML, JSON, and HTML"
author: "Christophe"
date: "March 15, 2016"
output: html_document
---

```{r, pre load data, include = FALSE}
library(XML)
library(jsonlite)
library(df2json)
library(htmlTable)
library(knitr)
library(data.table)
library(RJSONIO)
url <- "https://raw.githubusercontent.com/ChristopheHunt/MSDA---Coursework/master/Data%20607/Homework/data/"
```

#### We load the data below
```{r}
books.url <- file(paste(url,"books.csv", sep = ""), open="r" )
books.data <- read.csv2(books.url, sep = ",", header=TRUE, stringsAsFactors = FALSE ,  encoding="UTF-8")
```

#### This is how the `books.data` from csv looks like
```{r}
books.data
```

#### The following code will take our `books.data` file and then convert it to `xml` format. 

```{r, warning=FALSE}
library(XML)
create_xml_books <- function(dataframe){
                      for (i in seq(1:length(dataframe[,1]))) {
                            xml$addTag(enc2utf8("book"), close=FALSE)
                           for (j in colnames(dataframe)) {
                              xml$addTag(j, dataframe[i,j])
                           } 
                              xml$closeTag()
                       }
                     }

xml <- xmlOutputDOM()
xml$addTag("books", close=FALSE)
create_xml_books(books.data)
xml$closeTag()

books.xml <- saveXML(xml$value()) 
```

#### Now lets look at the structure of the `xml` file. 

```{r}
cat(books.xml)
```

#### Now that we have the data in `xml` we can covert to a dataframe.

```{r}
books.xml.list <- xmlToList(books.xml)
books.xml.df <- as.data.frame(rbindlist(books.xml.list, fill = TRUE))
row.names(books.xml.df) <- names(books.xml.list$books$book)
kable(books.xml.df)
```

#### Lets load the dataframe to the `json` format using the package `df2json`

```{r, results = 'asis'}
books.json <- df2json::df2json(books.data)
```

#### Now we have a `json` format for the books data frame

```{r, results = 'asis'}
cat(books.json)
```

#### Lets now convert the json format back to a data frame using the json package `RJSONIO`

```{r}
books.json.list <- RJSONIO::fromJSON(books.json)
books.json.df <- as.data.frame(rbindlist(books.json.list, fill = TRUE))
rownames(books.json.df) <- NULL
kable(books.json.df)
```

#### Create the HTML file 

```{r}
books.html <- htmlTable(books.data)
```

```{r}
books.html
```

#### Now lets convert back our html file to a data frame using `XML` package

```{r}
books.html.df <- as.data.frame(XML::readHTMLTable(books.html))
column.names <- gsub("NULL.", replacement = "", colnames(books.html.df))
colnames(books.html.df) <- column.names
books.html.df <- books.html.df[,-c(1)]
kable(books.html.df)
```

#### Overall, I noticed that many of the results were returned in a list format. It seems that when parsing various data format types they results will be in a list format that needs to be modified into the data frame. The packages are extremely useful in producing these results. I would say that the `xml` required the most amount of effort in coding and understanding, whereas `json` felt to be the easiest to understand. I had not previously worked with data were the encoding to `UTF-8` was so critical to being able to parse the data, I now have a new understanding of this encoding. 
