---
title: "Distribution for Discussion - Data 604"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r, include = FALSE}
library(tidyverse)
library(rbokeh)
```


```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(rbokeh)

list_x <- NULL
for(i in 1:1000){
  x <- rgamma(n = 30, shape = 5)
  list_x[[i]] <- list(values = x, mean = mean(x), min = min(x))
}


figure(width = 600, height = 400, legend_location = "top_left") %>%
  ly_hist(list_x %>% map_dbl("mean"), breaks = 40, freq = FALSE, color = "blue") %>%
  ly_hist(list_x %>% map_dbl("min") , breaks = 40, freq = FALSE, color = "red") %>%
  x_axis(label = "Value") %>% 
  theme_legend( background_fill_alpha = 0.95)
```

> <span style="color:red">Red</span> = Minimum, <span style="color:blue">Blue</span> = Mean


For the first week, we have a simple warm-up exercise for the discussion.  Using R, generate 100 simulations of 30 samples each from a distribution (other than normal) of your choice.  Graph the sampling distribution of means.  Graph the sampling distribution of the minimum. Share your graphs and your R code. What did the simulation of the means demonstrate?  What about the distribution of the minimum...?  

(If you are unfamiliar with generating random variates in R, Google can help...  rexp(), runif(), rpoi(),...)