---
title: "Homework 10"
author: "Christophe Hunt"
date: "April 3, 2017"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
always_allow_html: yes
---

```{r, include = FALSE}
library(tidyverse)
```

\newpage

# Page 469: problem 3

The following data were obtained for the growth of a sheep population introduced into a new environment on the island of Tasmania (adapted from J. Davidson "On the Growth of the Sheep of Tasmania" Trans. R. Soc. S. Australia.)

\begin{table}[!h]
\centering
\caption{My caption}
\label{my-label}
\begin{tabular}{l|llllll}
t (year) & 1814 & 1824 & 1834 & 1844 & 1854 & 1864 \\ \hline
P (t) & 125 & 275 & 830 & 1200 & 1750 & 1650
\end{tabular}
\end{table}

## a. Make an estimate of $M$ by graphing $P(t)$.

```{r, cache=TRUE}
t <- c(1814, 1824, 1834, 1844, 1854, 1864)
P.t <- c(125, 275, 830, 1200, 1750, 1650)
df <- as.data.frame(cbind(t, P.t))
```

```{r}
library(tidyverse)
number_ticks <- function(n) {function(limits) pretty(limits, n)} # http://stackoverflow.com/a/17257422/4153261
ggplot(df, aes(x = t, y = P.t)) +
  geom_point() +
  scale_y_continuous(breaks = number_ticks(15), limits = c(0, 2000)) + 
  xlim(1800, 1900) 
```

> Based on the above graph I would estimate $M$ = 2000. 

## b. Plot ln[P/(M-P)] against $t$. If a logistic curve seems reasonable, estimate $rM$ and $t$*.  

```{r}

```


# Page 478: problem 6

Suggest other phenomena for which the model described in the text might be used

# Page 481: problem 1

## a. Using the estimate that $d_a$ = 0.054$v^2$, where 0.054 has dimensions ft*hr^2/Mi^2, show that the constant $k$ in Equation (11.29) has the value 19.9 ft/sec^2. 

## b. Using the data in Table 4.4, plot $d_b$ in ft versuse v^2/2 in ft^2/sec^2 to estimate 1/$k$ directly. 

# Page 522: problem 21

Oxygen flows through on tube into a liter flasks filled with air, and the mixture of oxygen and air (considered well stirred) escapes through another tube. Assuming that air contains 21% oxygen, what percentage of oxygen will the flask contain after 5 L have passed through the intake tube? 

# Page 522: problem 22

If the average person breathes 20 times per minute, exhaling each time 100 in^3 of air containing 4% carbon dioxide. Find the percentage of carbon dioxide in the air of a 10,000 ft^3 closed room 1 hr after a class of 30 students enters. Assume that the air is fresh at the start, that the ventilators admit 1000 ft^3 of fresh air per minute, and that the fresh air contains 0.04% carbon dioxide. 