---
title: "Homework 5"
author: "Christophe Hunt"
date: "March 4, 2017"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

# Problem Set 1 

Consider the unsolvable system Ax = b as given below:
\[
\begin{bmatrix}
    1 & 0  \\
    1 & 1 \\
    1 & 3 \\
    1 & 4 \\
\end{bmatrix}
\begin{bmatrix}
x_1 \\
x_2 \\
\end{bmatrix}
=
\begin{bmatrix}
0 \\
8 \\
8 \\
20  \\
\end{bmatrix}
\]

## Write R Markdown script to compute $A^TA$ and $A^Tb$

```{r}
A <- matrix(c(1,1,1,1,0,1,3,4), ncol = 2)
b <- matrix(c(0,8,8,20))

ATA <- t(A) %*% A
ATb <- t(A) %*% b

results <- list("ATA" = ATA, "ATb" = ATb)
results
```



## Solve for $\hat{x}$ in R using the above computed matrices

```{r}
x <- solve(ATA) %*% ATb
x
```


## What is the squared error of this solution?

```{r}
p <- A %*% x
#b = p + e or e = p - b which we can substitute in our given values. 
e <- p - b
# we then sum the square of errors. 
e2 <- sum(e^2)
e2
```


## Find the exact solution with p instead of b

```{r}
options(scipen = 999)
p   <- matrix(c(1,5,13,17))
ATp <- t(A) %*% p
xp  <- solve(ATA) %*% ATp
p2  <- A %*% xp
e   <- p2-p
e
```

Essentially, the error vector e is = 0. 

```{r}
e2p <- sum(e^2)
e2p
```

Show that the error $e = b - p = [-1;3;-5;3]$.
```{r}
b - p
```

Show that the error $e$ is orthogonal to $p$ and to each of the columns of $A$. 

As per the week 5 handout - We know that when two vectors are orthogonal, their dot product is zero. 

```{r}
e*p
```

```{r}
sum(e*A[,1])
```


\newpage

# Problem Set 2 

Write an R markdown script that takes in the auto-mpg data, extracts an A matrix from the first 4 columns and b vector from the fifth (mpg) column.

Apparently, an added column of 1 is necessary to obtain an intercept. 

```{r}
x <- as.matrix(read.table("https://raw.githubusercontent.com/ChristopheHunt/MSDA---Coursework/master/Data%20605/Assignment%205/auto-mpg.data"))

A <- as.matrix(cbind(x[,1:4],1))
b <- as.matrix(x[,5])
```

Using the least squares approach, your code should compute the best fitting solution

```{r}
ATA <- t(A) %*% A
ATb <- t(A) %*% b
results <- list("ATA" = ATA, "ATb" = ATb)
results
```

\newpage

## Solve for $\hat{x}$ in R using the above computed matrices

```{r}
x <- solve(ATA) %*% ATb
x
```

The least squares model using this method is:

$$mpg = `r round(x[1,], 5)`*displacement + `r round(x[2,], 5)`*horsepower + `r round(x[3,], 5)`*weight + `r round(x[4,], 5)`*acceleration 
 + `r round(x[5,], 5)`$$


Finally, calculate the fitting error between the predicted mpg of youur model and actual mpg. 

## What is the squared error of this solution?

```{r}
p <- A %*% x
#b = p + e or e = p - b which we can substitute in our given values. 
e <- p - b
# we then sum the square of errors. 
e2 <- sum(e^2)
e2
```
