---
title: "Homework 5"
author: "Christophe Hunt"
date: "March 4, 2017"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
---

# Problem Set 1 

Consider the unsolvable system Ax = b as given below:
\[
\begin{bmatrix}
    1 & 0  \\
    1 & 1 \\
    1 & 3 \\
    1 & 4 \\
\end{bmatrix}
\begin{bmatrix}
x_1 \\
x_2 \\
\end{bmatrix}
=
\begin{bmatrix}
0 \\
8 \\
8 \\
20  \\
\end{bmatrix}
\]

## Write R Markdown script to compute $A^TA$ and $A^Tb$

```{r}
A <- matrix(c(1,1,1,1,0,1,3,4), ncol = 2)
b <- matrix(c(0,8,8,20))

ATA <- t(A) %*% A
ATb <- t(A) %*% b

results <- list("ATA" = ATA, "ATb" = ATb)
results
```



## Solve for $\hat{x}$ in R using the above computed matrices

```{r}
x <- solve(ATA) %*% ATb
x
```


## What is the squared error of this solution?

```{r}
p <- A %*% x
#b = p + e or e = p - b which we can substitute in our given values. 
e <- p - b
# we then sum the square of errors. 
e2 <- sum(e^2)
e2
```


## Find the exact solution with p instead of b

```{r}
options(scipen = 999)
p   <- matrix(c(1,5,13,17))
ATp <- t(A) %*% p
xp  <- solve(ATA) %*% ATp
p2  <- A %*% xp
e   <- p2-p
e
```

Essentially, the error vector e is = 0. 

```{r}
e2p <- sum(e^2)
e2p
```

Show that the error $e = b - p = [-1;3;-5;3]$.
```{r}
b - p
```

Show that the error $e$ is orthogonal to $p$ and to each of the columns of $A$. 

As per the week 5 handout - We know that when two vectors are orthogonal, their dot product is zero. 

```{r}
e*p
```

```{r}
sum(e*A[,1])
```


# Problem Set 2 
