---
title: "Week 2 - Subsetting Data w/R"
author: "Christophe"
date: "February 7, 2016"
output: html_document
---
```{r pre-load library, include=FALSE}
library(plyr)
library(htmlTable)
```
 Data Set Title: Flag database

Source Information:  
(a) Creators: Collected primarily from the "Collins Gem Guide to Flags": Collins Publishers (1986).  
(b) Donor: Richard S. Forsyth, 8 Grosvenor Avenue, Mapperley Park, Nottingham NG3 5DX, 0602-621676  
(c) Date: 5/15/1990

In the following R markdown I am converting a subset of the Flag Dataset from variables to their full attribute information. 

The actual data set is found here: http://archive.ics.uci.edu/ml/machine-learning-databases/flags/flag.data

I chose the first 3 columns and the 6th and column from the data set and transform the information based on the data dictionary here: http://archive.ics.uci.edu/ml/machine-learning-databases/flags/flag.names. 

 1. name: Name of the country concerned
 2. landmass:	1=N.America, 2=S.America, 3=Europe, 4=Africa, 5=Asia, 6=Oceania
 3. zone:	Geographic quadrant, based on Greenwich and the Equator 1=NE, 2=SE, 3=SW, 4=NW
 6. language: 1=English, 2=Spanish, 3=French, 4=German, 5=Slavic, 6=Other Indo-European, 7=Chinese, 8=Arabic, 9=Japanese/Turkish/Finnish/Magyar, 10=Others
 8. bars:    Number of vertical bars in the flag
 9. stripes:  Number of horizontal stripes in the flag
 10. colours:  Number of different colours in the flag

```{r}
library(plyr)
f.data <- file("http://archive.ics.uci.edu/ml/machine-learning-databases/flags/flag.data",
               open="r" )
data <- read.table(f.data, sep=",", header=FALSE, stringsAsFactors = FALSE)
subset <- data[, c( 1,2,3,6,8,9,10)]
colnames(subset) <- c("name","landmass","zone","language","bars","stripes","colours")
subset$landmass <- as.factor(mapvalues(subset$landmass, c(seq(1:6)), c("N.America","S.America","Europe","Africa","Asia","Oceania")))
subset$zone <- as.factor(mapvalues(subset$zone, c(seq(1:4)),c("NE","SE","SW","NW")))
subset$language <- as.factor(mapvalues(subset$language, c(seq(1:10)),c("English", "Spanish", "French", "German", "Slavic", "Other Indo-European", "Chinese", "Arabic", "Japanese/Turkish/Finnish/Magyar", "Others")))
close(f.data)

```

```{r}
library(htmlTable)
summary(subset)
htmlTable(head(subset))
```
