---
title: "Project 2"
author: "Christophe"
date: "March 12, 2016"
output: html_document
---
```{r pre-load, include = FALSE}
library(plyr)
library(sqldf)
library(dplyr)
library(scales)
library(tcltk)
library(tidyr)
library(knitr)
library(ggplot2)
url <- "https://raw.githubusercontent.com/ChristopheHunt/MSDA---Coursework/master/Data%20607/Homework/data/"
```

##Calories from Restuarants

Jeffrey Nieman proposed some interesting questions related to calorie intake of those that eat out compared to those that do not. I find the subject very interesting as well so I chose his data set to inspect. 

Jeffrey extracted the data from this [page](http://www.ars.usda.gov/SP2UserFiles/Place/80400530/pdf/1112/Table_53_RST_GEN_11.pdf) and provided a nice csv file. Many thanks to Jeffrey. 

```{r}
cal.url <- file(paste(url,"calories and restaurants - Jeffrey Nieman.csv", sep = ""), open="r" )
cal.intake <- read.csv(cal.url, sep=",", header=TRUE, stringsAsFactors = FALSE)
cal.intake$Gender[cal.intake$Gender == ""]  <- NA
cal.intake <-  cal.intake %>% 
               fill(Gender) %>% 
               mutate(percent.of.cals.from.rest = 
                      (intake.kcal.from.restaurants.for.all.individuals/ 
                         Total.Intake.kcal.for.all.individuals)) %>% 
               select(Gender, Age, percent.of.cals.from.rest)

cal.intake$Age <- as.factor(cal.intake$Age)
levels(cal.intake$Age) <- unique(cal.intake$Age)

ggplot(cal.intake, aes(x = Age, y = percent.of.cals.from.rest, fill = Gender)) +
       ylab("Percent of Overall Calories from Restuarants\n") +
      xlab("\nAge Group") + 
      scale_y_continuous(label = percent) + 
      ggtitle("Percentage of Calories obtained from Restuarants by Age Group and Gender\n")+
      geom_bar(stat="identity", position = "dodge", width = .7)

colnames(cal.intake)[3] <- "Percent of Calories obtained from Restuarants"
cal.intake[,3] <- percent(cal.intake[,3])
kable(cal.intake, align = "c")
```

The results of the data is very interesting, I am surprised at the calories obtained from restuarants for the `2 to 5` age group. I wouldn't have thought that this group would be obtaining `r percent(cal.intake$percent.of.cals.from.rest[cal.intake$Gender == "Female" & cal.intake$Age == "2 to 5"])` for males and `r percent(cal.intake$percent.of.cals.from.rest[cal.intake$Gender == "Male" & cal.intake$Age == "2 to 5"])` for females. I would be interested to find out more about this rate. 

```{r}

```
