---
title: "Homework 13"
author: "Christophe Hunt"
date: "May 6, 2017"
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    includes:
      in_header: header.tex
    number_sections: yes
    toc: yes
    toc_depth: 2
  html_document:
    toc: yes
    toc_depth: '2'
always_allow_html: yes
---

\newpage

# Write a program to compute the derivative of $f(x) = x^3 + 2x^2$ at any value of $x$. 

```{r}
deriv_limit <- function(func, x, h){
               f <- function(x) {eval(parse(text = func))}
               return((f(x + h) - f(x)) / h)
               }

deriv_limit(func = ('x^3 + 2*x^2'), x = 2, h = 0.000001)
deriv_limit(func = ('x^3 + 2*x^2'), x = 20, h = 0.000001)
```

Test using the analytic form

```{r, cache=TRUE}
deriv_analytic <- function(func, val, var){
                      f_x <- D(parse(text = func), var) 
                      assign(var, val)
                      return(eval(f_x))
                    }
deriv_analytic(func = 'x^3 + 2*x^2', val = 2, var = 'x')
deriv_analytic(func = 'x^3 + 2*x^2', val = 20, var = 'x')
```

Your function should take in a value of x and return back an approximation to the derivative of $f(x)$ evaluated at that value. You should not use the analytical form of the derivative to compute it. Instead, you should compute this approximation using limits.

# Now, write a program to compute the area under the curve for the function $3x^2+4x$ in the range x = $[1, 3]$. 

```{r}
auc <- function(func, range = seq(from = 1, to = 3, by = 0.000001)){
        return(sum((function(x) {eval(parse(text = func))})(range) * 0.000001))
       }
auc(func = '3*x^2+4*x')
```

You should first split the range into many small intervals using some really small $\delta x$ value (say 1e-6) and then compute the approximation to the area under the curve.

# Please solve these problems analytically (i.e. by working out the math) and submit your answers.

## Use integration by parts to solve for Rsin(x)cos(x)dx 

## Use integration by parts to solve for Rx2exdx

## What is d dx (x cos(x)