---
title: "HA 6 (HW 6.2)"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r, include=FALSE}
library(fma)
library(ggfortify)
library(tidyverse)
library(psych)
```

2. The data below represent the monthly sales (in thousands) of product A for a plastics manufacturer for years 1 through 5 (data set `plastics`). 

```{r}
data(plastics)
```

a. Plot the time series of sales of product A. Can you identify seasonal fluctuations and/or a trend?

b. Use a classical multiplicative decomposition to calculate the trend-cycle and seasonal indices.

c. Do the results support the graphical interpretation from part (a)?

d. Compute and plot the seasonally adjusted data.

e. Change one observation to be an outlier (e.g., add 500 to one observation), and recompute the seasonally adjusted data. What is the effect of the outlier?

f. Does it make any difference if the outlier is near the end rather than in the middle of the time series?

g. Use a random walk with drift to produce forecasts of the seasonally adjusted data.

h. Reseasonalize the results to give forecasts on the original scale. 